Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-20 Manuscript/manuscript.Rmd
wip

Replicate Hi-C data for the GM12878 cell line were downloaded from GEO GSE63525 (Experiments HIC001 to HIC018; produced using the MbolI restriction enzyme) [@rao20143d]. The autosomal genomic coordinates (GRCh37/hg19 human genome assembly) of TAD boundaries were obtained from data at 10 kb, 25 kb, 50 kb, and 100 kb resolution using the Arrowhead algorithm [@durand2016juicebox]. The start and end coordinates of all TADs were concatenated, sorted, and unique TAD boundary points were obtained.

@@

The response vector $Y$ was formed by genomic regions defined by the resolution of the data. Genomic regions flanking TAD boundaries were labeled as $Y=1$; otherwise, the regions were labeled as not containing a boundary ($Y=0$).

## GENOMIC ANNOTATION DATA

Genomic annotation data (GRCh37/hg19 human genome assembly) were obtained from the UCSC Genome Browser Database [@Tyner:2017aa] (Supplementary Table 1). The annotations consisted of a variety of different chromatin states, histone marks, and transcription factor binding sites using ChIP-seq peak experiments. Each genomic annotation was represented by a set of regions (their genomic coordinates) annotated as having a functional property or biological significance.

@@

### FEATURE ENGINEERING AND REDUCTION USING GENOMIC ANNOTATIONS

Using the sets of genomic annotations, we created three types of predictors that were used as features to analyze their relationships with TAD boundaries (Figure 3):

1. **Overlap Counts (OC)**: For each genomic bin, the total number of instances where a bin overlapped with genomic annotation regions (features) was calculated. In the case of no overlaps the value is set as 0.

2. **Overlap Percent (OP)**: For each genomic bin, the percent overlap between the feature width and the total width of the bin was calculated. We defined the feature width as the number of bases that overlapped with the genomic bin. The percent was then calculated by dividing the feature width by the bin width. If multiple overlaps existed, the feature width was defined as the sum of the total number of bases that overlapped with a particular bin. Note that the bin width remained constant for a given resolution (either 10 kb, 25 kb, 50 kb, or 100 kb). Values for bins with no overlaps were set to be 0.

3. **Log Distance**: For each genomic bin, the log of the distance (in bases) from the center of the bin to the center of the nearest region of a respective genomic annotation region was calculated. The log transformation was used as a normalization technique in order to account for the skewness of the data (Supplementary Figure 1)

>>>>

Replicate Hi-C data for the GM12878 cell line were downloaded from GEO GSE63525 (Experiments HIC001 to HIC018; produced using the MbolI restriction enzyme) [@rao20143d]. The autosomal genomic coordinates (GRCh37/hg19 human genome assembly) of TAD boundaries were obtained from data at 10 kb, 25 kb, 50 kb, and 100 kb resolution using the Arrowhead algorithm [@durand2016juicebox]. The start and end coordinates of all TADs were concatenated, sorted, and unique TAD boundary points were obtained. Genomic annotation data (GRCh37/hg19 human genome assembly) were obtained from the UCSC Genome Browser Database [@Tyner:2017aa] (Supplementary Table S1).

@@

The response vector $Y$ was formed by genomic regions (bins) defined by the resolution of the data. Genomic regions flanking TAD boundaries were labeled as $Y=1$; otherwise, the regions were labeled as not containing a boundary ($Y=0$).

Using the sets of genomic features, we created three types of predictors (Figure 2):

1. **Overlap Counts (OC)**: For each genomic bin, the total number of overlaps with genomic features was calculated. In the case of no overlaps the value is set as 0.

2. **Overlap Percent (OP)**: For each genomic bin, the percent overlap was defined as the total width of features overlapping a bin divided by the bin width. Values for bins with no overlaps were set to be 0.

3. **Distance**: For each genomic bin, the log [??? didn't you test raw distance first?] [??? log10?] of the distance (in bases) from the center of the bin to the center of the nearest genomic feature region was calculated. The log transformation was used as a normalization technique in order to account for the skewness of the data (Supplementary Figure 1)

@@
